#Reproducible Research Course Project 1
By David English

##Initialisation
The first step is to ensure the correct libraraies are loaded. These are loaded using the following code.

```{r results='hide', warning=FALSE, message=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(lubridate)
```

##Loading the Data
Create a data table named *activity* that contains the data in the activity.csv file. The date column is subsequently transformed into a date format, rather than a character string.
``` {r}
activity <- fread("activity.csv") %>%
      mutate(date = ymd(date))
```

##Number of Steps Per Day
The data should be grouped by the date and then the data can be summarised to give the number of steps per day in a new data table.

```{r warning=FALSE, message=FALSE}
activity <- group_by(activity, date)
activity_sums <- summarise(activity, sum(steps)) %>%
                  rename(sums = `sum(steps)`)
ggplot(activity_sums, aes(x = date, y = sums)) + 
      geom_histogram(bins = length(activity_sums$date), stat = "identity")
```
